---
author: Alma Garcia
date: 30/OCT/2023
---

# Assignment B1: Making a function

```{r}
library(tidyverse)
library(testthat)
```

## Defining the function
I will create a function for getting the first n observations of a variable in a tibble. This wraps the dpyrl functions `slice_head` and `pull`, which I used for my MDA milestone 2. The function basically just requires the `dplyr` package. 

```{r}
#' Get top observations of a variable in a tibble
#' 
#' `get_observations` returns a list of the first `n` observations of variable `var` in tibble `.data`. In contrast to using `$` and `[n]`, `get_observations` can be piped. 
#' 
#' @param .data A data frame, data frame extension (e.g. a tibble), or a lazy data frame (e.g. from dbplyr or dtplyr). Named the same as the arguments from `slice_head()` and `pull()`.
#' @param var A variable specified as:
#' - a literal variable name
#' - a positive integer, giving the position counting from the left
#' - a negative integer, giving the position counting from the right
#' The default returns the first column. Named the same as the arguments from `slice_head()` and `pull()`.
#' @param n The number of rows to select. Default is `n` = 1. If `n` is greater than the number of rows in `.data` the result will be silently truncated to the `.data` row size. A negative value of `n` will be subtracted from the `.data` size. For example, `n = -2` with a `.data` of 5 rows will select 5 - 2 = 3 rows. Named the same as the arguments from `slice_head()` and `pull()`.
#' @param na_rm Determines whether to keep or drop rows with `NA.` Default is `FALSE`. I used this name because it is usually called like that in other functions. 
#' @returns A vector the same size as `.data`.
get_observations <- function(.data, var = 1, n = 1, na_rm = FALSE) {
    # stops and gives error message if .data is not the correct input
    if(!is.data.frame(.data) && !is_tibble(.data)) {
      stop('This function only works for dataframes or tibbles as input!\n',
         'You provided an object of class: ', class(.data)[1])
    }
    
  # stops and gives error message if n = 0
    if(n == 0) {
      stop('n must be > 0 or < 0')
    }
  
    observations <- .data %>%
    select({{ var }}) %>%
    # deal with NAs
    {if(na_rm == TRUE) drop_na(.) else .} %>%
    slice_head(n = n) %>%
    pull()
    return(observations)
}
```


## Examples

Getting the first 10 observations from the `wt` column from `mtcars`
```{r}
mtcars %>% 
  get_observations(wt, 10)
```
Working with other functions (piping), in this case to sort `wt` in descending order and then getting the first observation. 

```{r}
mtcars %>%
  arrange(desc(wt)) %>%
  get_observations(wt)
```

Removing NAs

```{r}
as_tibble(c(1,2,NA,3,4)) %>%
  get_observations(n = 5, na_rm = TRUE)
```


## Testing
Here I show three different tests for the function. 

```{r}
test_that("get_observations works", {
  # test for different class of .data error
  expect_error(get_observations("Observation"), "This function only works for dataframes or tibbles as input!")
  # test for inputting n = 0 (if not stopped it will return numeric(0))
  expect_error(get_observations(mtcars, n = 0), "> 0 or < 0" )
  # test that it returns a vector of  equal length as the number of rows of the .data
  expect_equal(nrow(mtcars), length(get_observations(mtcars, n = 100)))
}
)
```

