---
author: Alma Garcia
date: 30/OCT/2023
---

# Assignment B1: Making a function

```{r}
library(tidyverse)
library(testthat)
```

## Defining the function
I will create a function for getting the first `n` observations of a variable in a dataframe or tibble; in other words, this function will extract the first `n` observations of a column into a vector. The function replaced the `dyplr` functions `slice_head` and `pull`, which I used for my MDA milestone 2 a few times. The function basically just requires the `dplyr` package. 

```{r}
#' Get top observations of a variable in a tibble
#' 
#' `get_observations` returns a list of the first `n` observations of variable `var` in tibble `.data`. In contrast to using `$` and `[n]`, `get_observations` can be piped. 
#' 
#' @param .data A data frame, data frame extension (e.g. a tibble), or a lazy data frame (e.g. from dbplyr or dtplyr). Named the same as the arguments from `slice_head()` and `pull()`.
#' @param var A variable specified as:
#' - a literal variable name
#' - a positive integer, giving the position counting from the left
#' - a negative integer, giving the position counting from the right
#' The default returns the first column. Named the same as the arguments from `slice_head()` and `pull()`.
#' @param n The number of rows to select. Default is `n` = 1. If `n` is greater than the number of rows in `.data` the result will be silently truncated to the `.data` row size. A negative value of `n` will drop the first `n` rows. For example, `n = -4` will return observations starting from row 5 until the last row.  Named the same as the arguments from `slice_head()` and `pull()`.
#' @param na_rm Determines whether to keep or drop rows with `NA.` Default is `FALSE`. I used this name because it is usually called like that in other functions. 
#' @returns A vector size `n` when `n` is positive; length of `.data` - `n` when `n` is negative; and length of `.data` when `n` exceeds the number of rows in data.
get_observations <- function(.data, var = 1, n = 1, na_rm = FALSE) {
    # stops and gives error message if .data is not the correct input
    if(!is.data.frame(.data) && !is_tibble(.data)) {
      stop('This function only works for dataframes or tibbles as input!\n',
         'You provided an object of class: ', class(.data)[1])
    }
  
    # stops and gives error message if n = 0
    if(n == 0) {
      stop('n must be > 0 or < 0')
    }
    
    # changes n to length of .data if n > nrow(.data)
    if(n > nrow(.data)) {
      n <- nrow(.data)
    }
  
    observations <- .data %>%
      {if(na_rm == TRUE) pull(.,{{ var }}) %>% na.omit(.) else pull(.,{{ var }})}
    
    return(observations[0:n])
}
```


## Examples

Getting the first 10 observations from the `wt` column from `mtcars`
```{r}
mtcars %>% 
  get_observations(wt, 10)
```
Working with other functions (piping), in this case to sort `wt` in descending order and then getting the first 5 observation. 

```{r}
mtcars %>%
  arrange(desc(wt)) %>%
  get_observations(wt, 5)
```

Removing NAs 

```{r}
mydata <- tibble(
  id   = 1:4,
  sex  = c("Male", "Female", "Female", "Male"),
  var1 = c(4, 1, 2, 3),
  var2 = c(NA, 4, 5, NA),
  var3 = c(2, 1, NA, NA)
)

mydata %>% get_observations(var2, n = 2, na_rm = TRUE)
```

Using negative numbers

```{r}
mtcars %>%
  get_observations(wt, -5)
```



## Testing
Here I show three different tests for the function. 

```{r}
test_that("get_observations works", {
  # test for different class of .data error
  expect_error(get_observations("Observation"), "This function only works for dataframes or tibbles as input!")
  # test for inputting n = 0 (if not stopped it will return numeric(0))
  expect_error(get_observations(mtcars, n = 0), "> 0 or < 0" )
  # test that it returns a vector of  equal length as the number of rows of the .data
  expect_equal(nrow(mtcars), length(get_observations(mtcars, n = 100)))
}
)
```

